---
title: "Get annotated, differentially-methylated regions by comparing samples from two conditions"
author: "Tsunghan Hsieh"
description: >
  This file demonstrates how to get annotated DMR.
date: "`r format(Sys.Date(), format='%d-%m-%Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Clustering_Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set RMarkDown options
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, eval = TRUE, collapse = FALSE)

# Turn off scientific notation
options(scipen = 999)

# load all functions
devtools::document()
devtools::load_all()
```

# Get adjusted methylation counts
***
We first get the methylation counts for each sample.

```{r todo}
# Run the following Rmd file:
# ./Doc/getUnadjustedMethylationCount2018.rmd.
# ./Doc/getUnadjustedMethylationCount2022.rmd.
```

# Set-up
***

## Load libraries
```{r library}
library(dplyr)
library(methylKit)
library(here)
library(stringr)
```

## Set dir
***
Set input and output dir.

```{r dir}
sampleSheetDir <- here::here("Doc","sampleSheet")
inputDir5mc <- here::here("Results", "Statistics", "5mc")
inputDir5hmc <- here::here("Results", "Statistics", "5hmc")
inputDirTotal <- here::here("Results", "Statistics", "5hmc_5mc")
outputDir <- here::here("Results", "Statistics", "DMR")
outputDirAnnotated <- here::here("Results", "Statistics", "DMR_annotated")
```


# Get differentially-methylated region (DMR) data
***
For each tissue and each methylation type, we compare samples from germ-free (GF) vs specific-pathogen-free (SPF) conditions. 5mc stands for 5'-methylcytosine, 5hmc stands for 5'-hydroxylmethylcotosine, and 5mc_5hmc (or total) stands for 5mc + 5hmc.

## 5mc
***


```{r sample_file}
## Check getDMR_2018.R
## Check getDMR_2022.R

## 5mc
sampleSheet <- data.frame(fname = list.files(inputDir5mc))

sampleSheet <- sampleSheet |>
  tidyr::separate(fname, c("tissue", "microbiome", "age", "ox", "replicate", "methylation", "format"), sep = "_|\\.")

## build getDMR.R

sampleSheet <- read.csv(file.path(sampleSheetDir, paste0("RRBS", YEAR,".csv")))
sampleSheet  <- sampleSheet [,2:ncol(sampleSheet)]

sampleSheetTotal <- sampleSheet |>
  dplyr::filter(OX == "-") |>
  dplyr::mutate(fpath = file.path(inputDirTotal,paste0(paste0(index,"_total.csv"))))

sampleSheet5hmc <- sampleSheet |>
  dplyr::filter(OX == "+") |>
  dplyr::mutate(fpath = file.path(outputTmpDir,paste0(paste0(index,"_5hmc.csv"))))

sampleSheet5mc <- sampleSheet |>
  dplyr::filter(OX == "+") |>
  dplyr::mutate(fpath = file.path(outputTmpDir,paste0(paste0(index,"_5mc.csv"))))

# Read methylation call list (tile = 100) ====
tile <- readRDS(file.path(inputTmpDir, paste0("RRBS_", YEAR, "_", TISSUE, "_methTile100_", ".rds")))

# Confirm the sample ID and sampleSheet
sampleIdList <- getSampleID(tile)
for (i in 1:length(sampleIdList)) {
  if (sampleIdList[i] == sampleSheet[i,]$index)
    print("TRUE")
}
```

# Get DMR
***
Get DMR for 5hmc + 5mc, 5hmc, and 5mc.

```{r DMR}

```

# Get annotated DMR
Get annotated DMR for 5hmc + 5mc, 5hmc, and 5mc.
```{r annotated_DMR}
dmrList <- list.files(outputDir)
for (file in dmrList) {
  annotateDMR(
    input = here::here(outputDir,file),
    col_names = c("chr", "start", "end", "strand"),
    output = NULL)
}
```
